SunPy - Python for Solar Physics

### 2014-08-27 1-Introduction.tex

A typical workflow begins with a solar physicist manually identifying
a small number of events of interest on the Sun.  This is typically
done in order to investigate in detail the physics of these events
(for example, the large solar flare of 23 July 2002 has Astrophysical
Journal Letters volume 595, dedicated to the analysis of this flare).
In this workflow, data from many different instruments are obtained.
These data are typically in many different storage formats (for
example, FITS (Flexible Image Transport System, \cite{refId0}), CSV, or
binary files), and contain many different types of data (such as
images, lightcurves and spectra).  The repositories these data reside
in can have different access methods.  This workflow is characterized
by the large number of heterogeneous datasets used in the
investigation of a small number of solar events.

>>>>

A typical workflow begins with a solar physicist manually identifying
a small number of events of interest on the Sun.  This is typically
done in order to investigate in detail the physics of these events
(for example, the large solar flare of 23 July 2002 has Astrophysical
Journal Letters volume 595, dedicated to the analysis of this flare).
In this workflow, data from many different instruments are obtained.
These data are typically in many different storage formats (for
example, FITS, CSV, or binary files), and contain many different types
of data (such as images, lightcurves and spectra).  The repositories
these data reside in can have different access methods.  This workflow
is characterized by the large number of heterogeneous datasets used in
the investigation of a small number of solar events.

### 2014-08-27 2-1-Map.tex

The architecture of the map subpackage consists of a template map called
\texttt{GenericMap}, which is a subclass of \texttt{astropy.nddata.NDData}. 
\texttt{NDData} is a generic wrapper around a \texttt{numpy.ndarray} with a 
\texttt{meta} attribute to store metadata.
As \texttt{NDData} is currently still in development, \texttt{GenericMap} does 
not yet make full use of its capabilities, but this inheritance structure 
provides for future integration with \texttt{astropy}. In order to provide 
instrument- or detector-specific integration, \texttt{GenericMap} is designed
to be subclassed. Each subclass of \texttt{GenericMap} can register 
with the \texttt{Map} creation factory, which will then automatically return an instance
of the specific \texttt{GenericMap} subclass dependent upon the data provided. 
SunPy v0.4 has \texttt{GenericMap} specialisations for the following 
instruments: 
\textit{Yohkoh}/SXT, \textit{SOHO}/EIT and LASCO, \textit{RHESSI}, 
\textit{STEREO}/EUVI and COR, \textit{Hinode}/XRT,
\textit{PROBA2}/SWAP, \textit{SDO}/AIA and HMI, 
and \textit{IRIS} SJI frames.

The \texttt{Map} class stores all of the metadata retrieved from the header of
the image file in the \texttt{meta} attribute and provides convenience 
properties for commonly accessed metadata: e.g., \texttt{instrument}, 
\texttt{wavelength} or \texttt{coordinate\_system}. 
Listing \ref{code:aia_1} demonstrates the quick-look functionality of 
\texttt{Map}.

>>>>

The architecture of the map subpackage consists of a template map called
\texttt{GenericMap}, which is a subclass of \texttt{astropy.nddata.NDData}. 
\texttt{NDData} is a generic wrapper around a \texttt{numpy.ndarray} with a 
\texttt{meta} attribute to store metadata.
As \texttt{NDData} is currently still in development, \texttt{GenericMap} does 
not yet make full use of its capabilities, but this inheritance structure 
provides for future integration with \texttt{astropy}. In order to provide 
instrument- or detector-specific integration, \texttt{GenericMap} is designed
to be subclassed. Each subclass of \texttt{GenericMap} can register 
with the \texttt{Map} creation factory, which will then automatically return an instance
of the specific \texttt{GenericMap} subclass dependent upon the data provided. 
SunPy v0.4 has \texttt{GenericMap} specialisations for the following 
instruments: 
\textit{Yohkoh}/SXT, \textit{SOHO}/EIT and LASCO, \textit{RHESSI}, 
\textit{STEREO}/EUVI and COR, \textit{Hinode}/XRT,
\textit{PROBA2}/SWAP, \textit{SDO}/AIA and HMI, 
and \textit{IRIS} SJI (slit-jaw imager) frames.

The \texttt{GenericMap} class stores all of the metadata retrieved from the header of
the image file in the \texttt{meta} attribute and provides convenience 
properties for commonly accessed metadata: e.g., \texttt{instrument}, 
\texttt{wavelength} or \texttt{coordinate\_system}.
These properties are dynamic mappings to the underlying metadata and all methods 
of the \textit{GenericMap} class modify the meta data where needed.
For example, if \verb|aiamap.meta[`instrume']| is modified then \verb|aiamap.instrument| 
will reflect this change.
Currently this is implemented by not preserving the keywords of the input data,
instead modifying meta data to a set of ``standard" keys supported by SunPy.
Listing \ref{code:aia_1} demonstrates the quick-look functionality of 
\texttt{Map}.

### 2014-08-27 3-4-Helioviewer.tex

SunPy provides the ability to download images hosted by the
Helioviewer Project (\href{http://helioviewer.org}{http://helioviewer.org}).  
The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogues) via
easy-to-use visual interfaces. The Helioviewer Project have developed two client 
applications that allow users to browse images and create movies of the Sun taken 
by a variety of instruments: \href{http://www.helioviewer.org}{Helioviewer.org}, a 
Google Maps-like web application, and \href{http://www.jhelioviewer.org}{JHelioviewer}, 
a movie streaming desktop application. In order to manage all of this data, the Helioviewer
project maintains archives of all of its data in JPEG2000 format (for further
details on the Helioviewer Project see \citealt{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files from which they are generated, but are still high-fidelity, and thus can be used to quickly
visualise large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

@@

\begin{listing}[H]
\pythoncode{helioviewer.py}
\begin{center}
\includegraphics[width=0.6\columnwidth]{helioviewer_overlay_example}
\end{center}
\caption{Acquisition of a PNG image composed from data from three
  separate sources.}
\label{code:hv:overlaid}
\end{listing}

@@

\begin{listing}[H]
\pythoncode{helioviewer2.py}
\begin{center}
\includegraphics[width=0.8\columnwidth]{helioviewer_hmi_continuum_jp2_to_map}
\end{center}
\caption{Acquisition and display of a Helioviewer JPEG2000 file as a
  SunPy \texttt{Map} object. Images values are byte-scaled in the range 0--255.}
\label{code:downloadjp2}
\end{listing}

>>>>

SunPy provides the ability to download images hosted by the
Helioviewer Project (\href{http://helioviewer.org}{http://helioviewer.org}).  
The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogues) via
easy-to-use visual interfaces. The Helioviewer Project have developed two client 
applications that allow users to browse images and create movies of the Sun taken 
by a variety of instruments: \href{http://www.helioviewer.org}{Helioviewer.org}, a 
Google Maps-like web application, and \href{http://www.jhelioviewer.org}{JHelioviewer}, 
a movie streaming desktop application. In order to manage all of this data, the Helioviewer
project maintains archives of all of its data in JPEG2000 format (for further
details on the Helioviewer Project, see \citealt{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files from which they are generated, but are still high-fidelity, and thus can be used to quickly
visualise large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

@@

\begin{listing}[H]
\pythoncode{pycode_helioviewer.txt}
\begin{center}
\includegraphics[width=0.6\columnwidth]{helioviewer_overlay_example}
\end{center}
\caption{Acquisition of a PNG image composed from data from three
  separate sources.}
\label{code:hv:overlaid}
\end{listing}

@@

\begin{listing}[H]
\pythoncode{pycode_helioviewer2.txt}
\begin{center}
\includegraphics[width=0.8\columnwidth]{helioviewer_hmi_continuum_jp2_to_map}
\end{center}
\caption{Acquisition and display of a Helioviewer JPEG2000 file as a
  SunPy \texttt{Map} object. Images values are byte-scaled in the range 0--255.}
\label{code:downloadjp2}
\end{listing}
